---
title: "Data Analysis of learning-phase data"
format: 
  html:
    embed-resources: true
    self-contained-math: true
    toc: true
    toc-depth: 6
    toc-location: left
editor: visual
---

### R preparation

```{r}
if (!requireNamespace("pacman", quietly = TRUE)) { 
  install.packages("pacman")
} 
library("pacman")
p_load(tidyverse, here, viridis, patchwork, ggtext, lme4)


df_learning <- read.csv("data_learning.csv")
```

***The goal of analyzing the learning-phase data is to evaluate the learning outcomes and prepare subsamples for robustness analysis.***

### Wiki

Cherry and banana are a pair of mutually exclusive items: cherry (coded as "A" in the judgment phase data) and banana (coded as "B" in the judgment phase data) are both from Box A, each having a generating probability of 0.6 and 0.4, respectively.

Orange and strawberry are a pair of mutually exclusive items: orange (coded as "C" in the judgment phase data) and strawberry (coded as "D" in the judgment phase data) are both from Box B, each having a generating probability of 0.5 and 0.5, respectively.

Carrot and tomato are a pair of mutually exclusive items: carrot and tomato (coded as "E" and "F" respectively in the judgment phase data) are both from Box C, each having a generating probability of 0.8 and 0.2 respectively.

Broccoli and radish are a pair of mutually exclusive items: broccoli and radish (coded as "G" and "H" respectively in the judgment phase data) are both from Box D, each having a generating probability of of 0.9 and 0.1 respectively.

We always coded/represented the two events of drawing each of one pair of mutually exclusive items using the notation A, not-A.

Strawberry vs Orange —\> A (0.5) v.s. not-A (0.5)

Banana vs Cherry —\> A (0.6) v.s. not-A (0.4)

Carrot versus Tomato —\> A (0.8) v.s. not-A (0.2)

Broccoli versus Radish —\> A (0.9) v.s. not-A (0.1)

### Dataset preparation

1\. Calculate the realized probabilities for event pairs within participant

```{r}
#The probability of banana and cherry being drawn realized in the experiment calculated within participants:
df_learning_banana <- df_learning |> 
  filter(button1 == "banana.svg" | button2 == "banana.svg") |> 
  select(where(~ !all(is.na(.)))) |> 
  group_by(id) |> 
  mutate(n_rows = n()) |> 
  group_by(id) |> 
  mutate(n_banana = sum(drawn_fruit == "banana.svg"),
         n_cherry = sum(drawn_fruit == "cherry.svg")) |> 
  ungroup() |> 
  mutate(prop_banana = n_banana / n_rows,
         prop_cherry = n_cherry / n_rows) |> 
  mutate(check = prop_banana + prop_cherry) |> 
  select(id, prop_banana, prop_cherry) |> 
  unique()

df_learning_banana |> 
  summarise(
    mean   = mean(prop_banana),
    sd     = sd(prop_banana),
    min    = min(prop_banana),
    max    = max(prop_banana),
    median = median(prop_banana)
  )

#The probability of orange and strawberry being drawn realized in the experiment calculated for each participant:
df_learning_orange <- df_learning |> 
  filter(button1 == "orange.svg" | button2 == "orange.svg") |> 
  select(where(~ !all(is.na(.)))) |> 
  group_by(id) |> 
  mutate(n_rows = n()) |> 
  group_by(id) |> 
  mutate(n_orange = sum(drawn_fruit == "orange.svg"),
         n_strawberry = sum(drawn_fruit == "strawberry.svg")) |> 
  ungroup() |> 
  mutate(prop_orange = n_orange / n_rows,
         prop_strawberry = n_strawberry / n_rows) |> 
  mutate(check = prop_orange + prop_strawberry) |> 
  select(id, prop_orange, prop_strawberry) |> 
  unique()

df_learning_orange |> 
  summarise(
    mean   = mean(prop_orange),
    sd     = sd(prop_orange),
    min    = min(prop_orange),
    max    = max(prop_orange),
    median = median(prop_orange)
  )


#The probability of broccoli and radish being drawn realized in the experiment calculated within participants:
df_learning_broccoli <- df_learning |> 
  filter(button1 == "broccoli.svg" | button2 == "broccoli.svg") |> 
  select(where(~ !all(is.na(.)))) |> 
  group_by(id) |> 
  mutate(n_rows = n()) |> 
  group_by(id) |> 
  mutate(n_broccoli = sum(drawn_veggie == "broccoli.svg"),
         n_radish  = sum(drawn_veggie == "raddish.svg")) |> 
  ungroup() |> 
  mutate(prop_broccoli = n_broccoli / n_rows,
         prop_radish = n_radish / n_rows) |> 
  mutate(check = prop_broccoli + prop_radish) |> 
  select(id, prop_broccoli, prop_radish) |> 
  unique()

df_learning_broccoli |> 
  summarise(
    mean   = mean(prop_broccoli),
    sd     = sd(prop_broccoli),
    min    = min(prop_broccoli),
    max    = max(prop_broccoli),
    median = median(prop_broccoli)
  )


#The probability of carrot and tomato being drawn realized in the experiment calculated within participants:
df_learning_carrot <- df_learning |> 
  filter(button1 == "carrot.svg" | button2 == "carrot.svg") |> 
  select(where(~ !all(is.na(.)))) |> 
  group_by(id) |> 
  mutate(n_rows = n()) |> 
  group_by(id) |> 
  mutate(n_carrot = sum(drawn_veggie == "carrot.svg"),
         n_tomato  = sum(drawn_veggie == "tomato.svg")) |> 
  ungroup() |> 
  mutate(prop_carrot = n_carrot / n_rows,
         prop_tomato = n_tomato / n_rows) |> 
  mutate(check = prop_carrot + prop_tomato) |> 
  select(id, prop_carrot, prop_tomato) |> 
  unique()

df_learning_carrot |> 
  summarise(
    mean   = mean(prop_carrot),
    sd     = sd(prop_carrot),
    min    = min(prop_carrot),
    max    = max(prop_carrot),
    median = median(prop_carrot)
      )

#merge the newly calculated "realized probabilities" dataset with the "df_learning" dataset
realized_prs <-  df_learning_broccoli |> 
  full_join(df_learning_carrot, by = "id") |> 
  full_join(df_learning_banana, by = "id") |> 
  full_join(df_learning_orange, by = "id") |> 
  select(id, realized_broccoli = prop_broccoli, realized_radish = prop_radish, 
         realized_carrot = prop_carrot, realized_tomato = prop_tomato,
         realized_banana = prop_banana, realized_cherry = prop_cherry,
         realized_orange = prop_orange, realized_strawberry = prop_strawberry) |> 
  mutate(br_box_optimal = case_when(realized_broccoli > realized_radish ~ "broccoli.svg",
                                    realized_broccoli < realized_radish ~ "raddish.svg",
                                    realized_broccoli == realized_radish ~ "tie",
                                    .unmatched = "error"),
         ct_box_optimal = case_when(realized_carrot > realized_tomato ~ "carrot.svg",
                                    realized_carrot < realized_tomato ~ "tomato.svg",
                                    realized_carrot == realized_tomato ~  "tie",
                                    .unmatched = "error"),
         bc_box_optimal = case_when(realized_banana > realized_cherry ~ "banana.svg",
                                    realized_banana < realized_cherry ~ "cherry.svg",
                                    realized_banana == realized_cherry ~ "tie",
                                    .unmatched = "error"),
         os_box_optimal = case_when(realized_orange > realized_strawberry ~ "orange.svg",
                                    realized_orange < realized_strawberry ~ "strawberry.svg",
                                    realized_orange == realized_strawberry ~ "tie",
                                    .unmatched = "error")
  ) 

df_learning <- full_join(df_learning, realized_prs, by = "id")
```

2.  Prepare dataset in the long format, rename the variables and variable values, and keep only the relevant variables

```{r}
#mutate variables indicating the trial type (which type of option pair being presented) and trial order 
df_learning <- df_learning |> 
  arrange(id, counter_learning_fruit)  |> 
  mutate(
    os_box  = ((button1 == "orange.svg" & button2 == "strawberry.svg") |
                  (button1 == "strawberry.svg" & button2 == "orange.svg")),
    bc_box = ((button1 == "banana.svg" & button2 == "cherry.svg") | 
      (button1 == "cherry.svg" & button2 == "banana.svg")),
    br_box = ((button1 == "broccoli.svg" & button2 == "raddish.svg")|
      (button1 == "raddish.svg" & button2 == "broccoli.svg")),
    ct_box = ((button1 == "carrot.svg" & button2 == "tomato.svg")|
                 (button1 == "tomato.svg" & button2 == "carrot.svg"))
    ) |> 
  filter(if_any(c(os_box, bc_box, br_box, ct_box), ~ !is.na(.x))) |>
  group_by(id)  |> 
  mutate(
    os_order = if_else(os_box, cumsum(os_box), NA),
    bc_order = if_else(bc_box, cumsum(bc_box), NA),
    br_order = if_else(br_box, cumsum(br_box), NA),
    ct_order = if_else(ct_box, cumsum(ct_box), NA)
  ) |> 
  mutate(
   os_phase = case_when(
     os_order %in% c(1:5) ~ 1,
     os_order %in% c(6:10) ~ 2,
     os_order %in% c(11:15) ~ 3,
     os_order %in% c(16:20) ~ 4,
     os_order %in% c(21:25) ~ 5,
     os_order %in% c(26:30) ~ 6,
     os_order %in% c(31:35) ~ 7,
     os_order %in% c(36:40) ~ 8
   ),
   bc_phase = case_when(
     bc_order %in% c(1:5) ~ 1,
     bc_order %in% c(6:10) ~ 2,
     bc_order %in% c(11:15) ~ 3,
     bc_order %in% c(16:20) ~ 4,
     bc_order %in% c(21:25) ~ 5,
     bc_order %in% c(26:30) ~ 6,
     bc_order %in% c(31:35) ~ 7,
     bc_order %in% c(36:40) ~ 8
   ),
   br_phase = case_when(
    br_order %in% c(1:5) ~ 1,
    br_order %in% c(6:10) ~ 2,
    br_order %in% c(11:15) ~ 3,
    br_order %in% c(16:20) ~ 4,
    br_order %in% c(21:25) ~ 5,
    br_order %in% c(26:30) ~ 6,
    br_order %in% c(31:35) ~ 7,
    br_order %in% c(36:40) ~ 8
   ),
   ct_phase = case_when(
    ct_order %in% c(1:5) ~ 1,
    ct_order %in% c(6:10) ~ 2,
    ct_order %in% c(11:15) ~ 3,
    ct_order %in% c(16:20) ~ 4,
    ct_order %in% c(21:25) ~ 5,
    ct_order %in% c(26:30) ~ 6,
    ct_order %in% c(31:35) ~ 7,
    ct_order %in% c(36:40) ~ 8 
   )
  ) |> 
  ungroup()

#mutate a variable calculating the proportion of optimal choice within participant
#mutate a variable calculating the proportion of A choice within participant
df_learning_summary <- df_learning |> 
  pivot_longer(
    cols = ends_with("_phase"),
    names_to = "pair_type",   # os_phase / bc_phase / ...
    values_to = "phase",
    values_drop_na = TRUE) |> 
  mutate(optimal_choice = case_when(
    pair_type == "os_phase" ~ os_box_optimal,
    pair_type == "bc_phase" ~ bc_box_optimal,
    pair_type == "br_phase" ~ br_box_optimal, 
    pair_type == "ct_phase" ~ ct_box_optimal,
    .unmatched = "error"
  )) |> 
  mutate(
    # nicer labels for facets
    pair_type = dplyr::recode(
      pair_type,
      "os_phase" = "A (0.50) vs not-A (0.50)",
      "bc_phase" = "A (0.60) vs not-A (0.40)",
      "ct_phase" = "A (0.80) vs not-A (0.20)",
      "br_phase" = "A (0.90) vs not-A (0.10)"
    )) |> 
  select(id, condition, button_chosen, pair_type, phase, optimal_choice,
         realized_A_0.90 = realized_broccoli, realized_A_0.80 = realized_carrot,
         realized_A_0.50 = realized_orange, realized_A_0.60 = realized_banana) |> 
  mutate(button_chosen = case_when(
    button_chosen %in% c("strawberry.svg", "banana.svg",
                         "carrot.svg", "broccoli.svg") ~ "A",
    button_chosen %in% c("orange.svg", "cherry.svg",
                         "tomato.svg", "raddish.svg" ) ~ "not-A",
    .unmatched = "error"
  )) |> 
  mutate(optimal_choice = case_when(
    optimal_choice %in% c("strawberry.svg", "banana.svg",
                          "carrot.svg", "broccoli.svg") ~ "A",
    optimal_choice %in% c("orange.svg", "cherry.svg",
                          "tomato.svg", "raddish.svg" ) ~ "not-A",
    optimal_choice == "tie" ~ "tie",
    .unmatched = "error"
  )) |> 
  mutate(if_A_chosen = case_when(
    button_chosen == "A" ~ 1,
    button_chosen == "not-A" ~ 0,
    .unmatched = "error"
  )) |> 
  mutate(
    if_optimal = ifelse(
      button_chosen == optimal_choice, 1, 0
    )) |>
  ungroup() |>
  group_by(id, condition, pair_type, phase) |> 
  mutate(prop_A = mean(if_A_chosen),
         prop_optimal = mean(if_optimal)) |>
  ungroup() |> 
  select(id, condition, pair_type, phase, optimal_choice, prop_A, prop_optimal,
         realized_A_0.90, realized_A_0.80, realized_A_0.60, realized_A_0.50,
         -if_A_chosen, -if_optimal, -button_chosen) |> 
  unique() 

pair_type_levels <- c(
  "A (0.50) vs not-A (0.50)",
  "A (0.60) vs not-A (0.40)",
  "A (0.80) vs not-A (0.20)",
  "A (0.90) vs not-A (0.10)"
)

df_learning_summary <- df_learning_summary  |> 
  mutate(pair_type = factor(pair_type, levels = pair_type_levels)) |> 
  mutate(condition = ifelse (condition %in% c("A", "B"), "low", "high")) 
```

### Plots

```{r}
italic_A <- as_labeller(function(x) gsub("\\bA\\b", "*A*", x))


condition_labels <- c("high" = "40 learning trials",
                      "low"  = "20 learning trials")

# distributions ----
df_dist <- df_learning_summary |> 
  distinct(id,
           `realized_A_0.50`, `realized_A_0.60`,
           `realized_A_0.80`, `realized_A_0.90`) |> 
  pivot_longer(starts_with("realized_A"),
               names_to = "pair_type", values_to = "Pr") |> 
  mutate(
    pair_type = recode(pair_type,
                       "realized_A_0.50" = "A (0.50) vs not-A (0.50)",
                       "realized_A_0.60" = "A (0.60) vs not-A (0.40)",
                       "realized_A_0.80" = "A (0.80) vs not-A (0.20)",
                       "realized_A_0.90" = "A (0.90) vs not-A (0.10)"),
    pair_type = factor(pair_type, levels = pair_type_levels)
  )

df_dist_means <- df_dist |> 
  group_by(pair_type) |> 
  summarise(Pr_mean = mean(Pr), .groups = "drop")

p_dist <- ggplot(df_dist, aes(x = Pr)) +
  geom_histogram(binwidth = 0.05, boundary = 0,
                 fill = "white", colour = "black") +
  geom_vline(data = df_dist_means, aes(xintercept = Pr_mean),
             colour = "#c1272d", linetype = "dashed", linewidth = 1) +
  geom_text(data = df_dist_means,
            aes(x = Pr_mean, y = Inf,
                label = sprintf("%.2f", Pr_mean)),
            colour = "#c1272d",
            hjust = -0.15, vjust = 1.6, size = 4) +
  facet_grid(. ~ pair_type, labeller = labeller(pair_type = italic_A)) +
  scale_x_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  coord_cartesian(xlim = c(0, 1)) +
  labs(x = "Probability of option *A*", y = "Number of participants") +
  theme_classic() +
  theme(strip.text = element_markdown(size = 12),
        plot.tag.position = c(0, 1.02))

# learning curves 
p_main <- ggplot(df_learning_summary, aes(x = phase, y = prop_A)) +
  geom_jitter(aes(colour = optimal_choice),
              position = position_jitter(width = 0.3, height = 0.04, seed = 1),
              shape = 1, size = 0.6, stroke = 0.3, alpha = 0.5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.4,
               linewidth = 0.4, colour = "black", aes(group = 1)) +
  stat_summary(fun = mean, geom = "line", linewidth = 0.5,
               colour = "black", aes(group = 1)) +
  stat_summary(fun = mean, geom = "point", size = 1.2,
               colour = "black", aes(group = 1)) +
  scale_colour_manual(
    name   = "Optimal choice",
    values = c("A" = "#D55E00", "not-A" = "#0072B2", "tie" = "#4D4D4D"),
    labels = c("A" = "*A*", "not-A" = "not-*A*", "tie" = "tie")) +
  guides(colour = guide_legend(
    override.aes = list(shape = 16, size = 3, alpha = 1))) +
  facet_grid(condition ~ pair_type,
             labeller = labeller(pair_type = italic_A,
                                 condition = condition_labels)) +
  scale_x_continuous(breaks = c(2, 4, 6, 8)) +
  labs(x = "Phase", y = "P(Option A chosen)") +
  theme_classic() +
  theme(
    strip.text.x = element_markdown(size = 12),   
    strip.text.y = element_markdown(size = 14)    
  )

# stack two plots
p_combined <- p_dist / p_main +
  plot_layout(heights = c(1, 3), guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(
    legend.position = "bottom",
    axis.title.x = element_markdown(size = 15),   # "Phase" + "Probability of option *A*"
    axis.title.y = element_text(size = 15),       # "P(Option A chosen)" + "Number of participants"
    legend.title = element_text(size = 14),
    legend.text  = element_markdown(size = 13),
    plot.tag     = element_text(size = 16, face = "bold")
  )

p_combined

ggsave("learning.pdf", width = 12, height = 8)
```

#### learning plot v2

```{r}
df_dist <- df_learning_summary |> 
  distinct(id,
           `realized_A_0.50`, `realized_A_0.60`,
           `realized_A_0.80`, `realized_A_0.90`) |> 
  pivot_longer(starts_with("realized_A"),
               names_to = "pair_type", values_to = "Pr") |> 
  mutate(
    pair_type = recode(pair_type,
                       "realized_A_0.50" = "A (0.50) vs not-A (0.50)",
                       "realized_A_0.60" = "A (0.60) vs not-A (0.40)",
                       "realized_A_0.80" = "A (0.80) vs not-A (0.20)",
                       "realized_A_0.90" = "A (0.90) vs not-A (0.10)"),
    pair_type = factor(pair_type, levels = pair_type_levels)
  )


df_dist_means <- df_dist |> 
  group_by(pair_type) |> 
  summarise(Pr_mean = mean(Pr), .groups = "drop")

p_dist <- ggplot(df_dist, aes(x = Pr)) +
  geom_histogram(binwidth = 0.05, boundary = 0,
                 fill = "white", colour = "black") +
  geom_vline(data = df_dist_means, aes(xintercept = Pr_mean),
             colour = "black", linetype = "dashed", linewidth = 0.5) +
  geom_text(data = df_dist_means,
            aes(x = Pr_mean, y = Inf,
                label = sprintf("%.2f", Pr_mean)),
            hjust = -0.15, vjust = 1.6, size = 4, colour = "red") +
  facet_grid(. ~ pair_type) +
  scale_x_continuous(breaks = c(0, 0.25,0.5, 0.75, 1)) +
  coord_cartesian(xlim = c(0, 1)) +
  labs(x = "Probability of option A", y = "Number of participants") +
  theme_classic() +
  theme(strip.text = element_markdown(size = 12),
        plot.tag.position = c(0, 1))  


# learning curves ----
p_main <- ggplot(df_learning_summary, aes(x = phase, y = prop_optimal)) +
  geom_jitter(aes(colour = optimal_choice),
              position = position_jitter(width = 0.3, height = 0.04, seed = 1),
              shape = 1, size = 0.6, stroke = 0.3, alpha = 0.5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.4,
               linewidth = 0.4, colour = "#c1272d", aes(group = 1)) +
  stat_summary(fun = mean, geom = "line", linewidth = 0.5,
               colour = "black", aes(group = 1)) +
  stat_summary(fun = mean, geom = "point", size = 1.2,
               colour = "black", aes(group = 1)) +
  scale_colour_manual(
    name   = "Optimal choice",
    values = c("A" = "#D55E00", "not-A" = "#0072B2", "tie" = "#4D4D4D")) +
  guides(colour = guide_legend(
    override.aes = list(shape = 16, size = 3, alpha = 1))) +
  facet_grid(condition ~ pair_type) +
  scale_x_continuous(breaks = c(2, 4, 6, 8)) +
  labs(x = "Phase", y = "P(Optimal option chosen)") +
  theme_classic() +
  theme(
    strip.text.x    = element_blank(),     # pair_type labels live on the top strip, not repeated here
    strip.text.y    = element_markdown(size=14),  # high / low on the right
    legend.position = "bottom"
  )

# ---- stack two plots, aligned by the pair_type columns, one shared legend ----
p_dist1 <- p_dist +
  theme(plot.tag.position = c(0, 1.02))

p_combined <- p_combined <- p_dist1 / p_main +
  plot_layout(heights = c(1, 3), guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "bottom") &
  theme(
    legend.position = "bottom",
    axis.title.x = element_text(size = 15),   # x-labs: "Block" + "Probability of option A"
    axis.title.y = element_text(size = 15),   # y-labs: "P(A option chosen)" + "Number of participants"
    legend.title = element_text(size = 14),
    legend.text  = element_text(size = 13),
    plot.tag     = element_text(size = 16, face = "bold")
  )  

p_combined

ggsave("learning_v2.pdf", width = 12, height = 8)
```

Learning outcomes were analyzed in small blocks (every 5 trials were grouped into one block), and were summarized with the probability of choosing the A option/proportion of A choices.

### Inferential Statistics

#### Stabilization

```{r}


```

#### Learning acquired or not?

```{r}

```

### Subsample based on learning outcomes

##### Try out exclusion criterion 1:

First, check what conditions the participant belongs to. When condition = “low", check what value the variable/column “prop_optimal” takes when “phase” = 4 and when pair types are edge paris. Only include the participants , when “pair_type” = edge types (i.e., pair_type = "A (0.90) vs not-A (0.10)" and pair_type = "A (0.80) vs not-A (0.20)" and “phase” = 4, “prop_optimal” \> 0.5. If participants belong to condition "high", check what value the variable/column “prop_optimal” takes when “phase” = 8 and when pair types are edge pairs. Only include the participant (ids) , when pair_type” = edge types and “phase” = 8, “prop_optimal” \> 0.5.

```{r}
# Qualifying participants 
included_ids <- df_learning_summary |> 
  # Keep only rows for edge pairs
  filter(pair_type %in% c("A (0.90) vs not-A (0.10)", "A (0.80) vs not-A (0.20)")) |> 
  # Keep only the relevant phase per condition
  filter(
    (condition == "low" & phase == 4) |
    (condition == "high" & phase == 8)
  ) |> 
  # For each participant & box combination, check if prop_optimal > 0.5
  group_by(id, pair_type) |> 
  summarise(passes = all(prop_optimal > 0.5)) |>
  ungroup() |> 
  # Each participant must pass BOTH edge boxes
  group_by(id) |> 
  mutate(include = sum(passes)) |> 
  filter(include == 2) #312 participants 

```

create subsample and save

```{r create subsample and save }
df_subsample_id <- included_ids |> 
  pull(id) |> unique()

saveRDS(df_subsample_id, "subsample_ids.rds")
```

##### Plotting sub-sample

```{r}
df_dist2 <- df_learning_summary |> 
  filter(id %in% df_subsample_id) |> 
  distinct(id,
           `realized_A_0.50`, `realized_A_0.60`,
           `realized_A_0.80`, `realized_A_0.90`) |> 
  pivot_longer(starts_with("realized_A"),
               names_to = "pair_type", values_to = "Pr") |> 
  mutate(
    pair_type = recode(pair_type,
                       "realized_A_0.50" = "A (0.50) vs not-A (0.50)",
                       "realized_A_0.60" = "A (0.60) vs not-A (0.40)",
                       "realized_A_0.80" = "A (0.80) vs not-A (0.20)",
                       "realized_A_0.90" = "A (0.90) vs not-A (0.10)"),
    pair_type = factor(pair_type, levels = pair_type_levels)
  )

df_dist2_means <- df_dist2 |> 
  group_by(pair_type) |> 
  summarise(Pr_mean = mean(Pr), .groups = "drop")

p_dist2 <- ggplot(df_dist2, aes(x = Pr)) +
  geom_histogram(binwidth = 0.05, boundary = 0,
                 fill = "#D55E00", colour = "white") +
  geom_vline(data = df_dist2_means, aes(xintercept = Pr_mean),
             colour = "black", linetype = "dashed", linewidth = 0.5) +
  geom_text(data = df_dist2_means,
            aes(x = Pr_mean, y = Inf,
                label = sprintf("%.2f", Pr_mean)),
            hjust = -0.15, vjust = 1.6, size = 4) +
  facet_grid(. ~ pair_type) +
  scale_x_continuous(breaks = c(0, 0.25,0.5, 0.75, 1)) +
  coord_cartesian(xlim = c(0, 1)) +
  labs(x = "Probability of option A", y = "Number of participants") +
  theme_classic() +
  theme(strip.text = element_markdown(size = 12),
        plot.tag.position = c(0, 1))  


# learning curves ----
df_learning_summary2 <- df_learning_summary |> 
  filter(id %in% df_subsample_id)

p_main2 <- ggplot(df_learning_summary2, aes(x = phase, y = prop_A)) +
  geom_jitter(aes(colour = optimal_choice),
              position = position_jitter(width = 0.3, height = 0.04, seed = 1),
              shape = 1, size = 0.6, stroke = 0.3, alpha = 0.5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.4,
               linewidth = 0.4, colour = "#c1272d", aes(group = 1)) +
  stat_summary(fun = mean, geom = "line", linewidth = 0.5,
               colour = "black", aes(group = 1)) +
  stat_summary(fun = mean, geom = "point", size = 1.2,
               colour = "black", aes(group = 1)) +
  scale_colour_manual(
    name   = "Optimal choice",
    values = c("A" = "#D55E00", "not-A" = "#0072B2", "tie" = "#4D4D4D")) +
  guides(colour = guide_legend(
    override.aes = list(shape = 16, size = 3, alpha = 1))) +
  facet_grid(condition ~ pair_type) +
  scale_x_continuous(breaks = c(2, 4, 6, 8)) +
  labs(x = "Phase", y = "P(Option A chosen)") +
  theme_classic() +
  theme(
    strip.text.x    = element_blank(),     # pair_type labels live on the top strip, not repeated here
    strip.text.y    = element_markdown(size=14),  # high / low on the right
    legend.position = "bottom"
  )

# ---- stack two plots, aligned by the pair_type columns, one shared legend ----
p1 <- p_dist2 +
  theme(plot.tag.position = c(0, 1.02))

p_combined <- p_combined <- p1 / p_main2 +
  plot_layout(heights = c(1, 3), guides = "collect") +
  plot_annotation(tag_levels = "A") &
  theme(legend.position = "bottom") &
  theme(
    legend.position = "bottom",
    axis.title.x = element_text(size = 15),   # x-labs: "Block" + "Probability of option A"
    axis.title.y = element_text(size = 15),   # y-labs: "P(A option chosen)" + "Number of participants"
    legend.title = element_text(size = 14),
    legend.text  = element_text(size = 13),
    plot.tag     = element_text(size = 16, face = "bold")
  )  

p_combined
```
