---
title: "Data Analysis"
format: 
  html:
    embed-resources: true
    self-contained-math: true
    toc: true
    toc-depth: 6
    toc-location: left
editor: visual
---

### R preparation

```{r}
if (!requireNamespace("pacman", quietly = TRUE)) { 
  install.packages("pacman")
} 
library("pacman")
#already load:
p_load(tidyverse, here, viridis, cowplot, MPTinR, scales, ggh4x, patchwork, irr)

df_judgment <- read.csv("data_judgment.csv") |> 
  mutate(frequency_condition = 
           case_when(condition %in% c("A", "B") ~ "20-learning-trial",
                     condition %in% c("C", "D") ~ "40-learning-trial",
                     TRUE ~ NA)) |>
  
 mutate(part = recode(part, "veggie" = "edge", "fruit" = "mid"))

subsample_ids <- readRDS("subsample_ids.rds")

```

### Wiki

Re-name the stimuli for the convenience of coding:

| banana (from box A): A
| cherry (from box A): B
| orange (form box B): C
| strawberry(from box B): D
| carrot(from box C): E
| tomato(from box C): F
| broccoli(from box D): G
| radish(from box D): H
| A&B are a pair of complementary mid-events, C&D are a pair of complementary mid-events,
| E&F are a pair of complementary edge-events, G&H are a pair of complementary edge-events

Liu et al. (2025) found illogical rankings. We wanted to know if illogical rankings occur after probability learning (in the learning environment). In the learning phase, participants learnt the probability of two pairs of complementary events {A, not-A, B, not-B}.Rankings were elicited from participants in two ways. In the first way, each participant provided rankings for stimuli of the same type (learned in the same block) simultaneously; that is, each participant provided a ranking for the mid-event set simultaneously, and then a ranking for the edge-event set simultaneously, resulting in two rankings per participant. In the second way, participants did pairwise comparisons of stimuli of the same type (learned in the same block). For each set of four stimuli of the same type, the participant did the pairwise comparison of every possible combination twice. For example, for the event set {A, not-A, B, not-B}, participants compared A versus not-A twice, compared A versus B twice, compared A versus not-B twice, compared not-A versus B twice, compared not-A versus not-B twice and compared B versus not-B twice. A ranking was elicited from the pairwise comparison results for the fruit set, and a ranking was elicited from the pairwise comparison results for the vegetable set, again resulting in two rankings per participant.

### Data preparation:

Specify the category system for categorizing the rankings. There are three types of rankings: logical rankings, interlaced illogical rankings, stacked illogical rankings and other illogical rankings.

```{r specifying cateogory system}

logical_patterns <- c("A\\|C\\|D\\|B",
                      "A\\|D\\|C\\|B",
                      "B\\|C\\|D\\|A",
                      "B\\|D\\|C\\|A",
                      "C\\|A\\|B\\|D",
                      "C\\|B\\|A\\|D",
                      "D\\|A\\|B\\|C",
                      "D\\|B\\|A\\|C",
                      "A\\|CD\\|B\\|",
                      "B\\|CD\\|A\\|",
                      "C\\|AB\\|D\\|",
                      "D\\|AB\\|C\\|",
                      "AC\\|BD\\|\\|",
                      "AD\\|BC\\|\\|",
                      "BC\\|AD\\|\\|",
                      "BD\\|AC\\|\\|",
                      "ABCD\\|\\|\\|",
                      "E\\|G\\|H\\|F",
                      "E\\|H\\|G\\|F",
                      "F\\|G\\|H\\|E",
                      "F\\|H\\|G\\|E",
                      "G\\|E\\|F\\|H",
                      "G\\|F\\|E\\|H",
                      "H\\|E\\|F\\|G",
                      "H\\|F\\|E\\|G",
                      "E\\|GH\\|F\\|",
                      "F\\|GH\\|E\\|",
                      "G\\|EF\\|H\\|",
                      "H\\|EF\\|G\\|",
                      "EG\\|FH\\|\\|",
                      "EH\\|FG\\|\\|",
                      "FG\\|EH\\|\\|",
                      "FH\\|EG\\|\\|",
                      "EFGH\\|\\|\\|")

stacked_patterns <- c("A\\|B\\|C\\|D", "A\\|B\\|D\\|C", 
                      "B\\|A\\|C\\|D", "B\\|A\\|D\\|C",
                      "C\\|D\\|A\\|B", "C\\|D\\|B\\|A",
                      "D\\|C\\|A\\|B", "D\\|C\\|B\\|A",
                      "AB\\|C\\|D\\|", "AB\\|D\\|C\\|",
                      "CD\\|A\\|B\\|", "CD\\|B\\|A\\|",
                      "A\\|B\\|CD\\|", "B\\|A\\|CD\\|",
                      "C\\|D\\|AB\\|", "D\\|C\\|AB\\|",
                      "AB\\|CD\\|\\|", "CD\\|AB\\|\\|",
                      "E\\|F\\|G\\|H", "E\\|F\\|H\\|G", 
                      "F\\|E\\|G\\|H", "F\\|E\\|H\\|G",
                      "G\\|H\\|E\\|F", "G\\|H\\|F\\|E",
                      "H\\|G\\|E\\|F", "H\\|G\\|F\\|E",
                      "EF\\|G\\|H\\|", "EF\\|H\\|G\\|",
                      "GH\\|E\\|F\\|", "GH\\|F\\|E\\|",
                      "E\\|F\\|GH\\|", "F\\|E\\|GH\\|",
                      "G\\|H\\|EF\\|", "H\\|G\\|EF\\|",
                      "EF\\|GH\\|\\|", "GH\\|EF\\|\\|")

interlaced_patterns <- c("A\\|C\\|B\\|D", "A\\|D\\|B\\|C",
                         "B\\|C\\|A\\|D", "B\\|D\\|A\\|C",
                         "C\\|A\\|D\\|B", "C\\|B\\|D\\|A",
                         "D\\|A\\|C\\|B", "D\\|B\\|C\\|A",
                         "AC\\|B\\|D\\|", "AC\\|D\\|B\\|",
                         "AD\\|B\\|C\\|", "AD\\|C\\|B\\|",
                         "BC\\|A\\|D\\|", "BC\\|D\\|A\\|",
                         "BD\\|A\\|C\\|",  "BD\\|C\\|A\\|",
                         "A\\|C\\|BD\\|",  "A\\|D\\|BC\\|",
                         "B\\|C\\|AD\\|",  "B\\|D\\|AC\\|",
                         "C\\|A\\|BD\\|",  "C\\|B\\|AD\\|",
                         "D\\|A\\|BC\\|",  "D\\|B\\|AC\\|",
                         "E\\|G\\|F\\|H", "E\\|H\\|F\\|G",
                         "F\\|G\\|E\\|H", "F\\|H\\|E\\|G",
                         "G\\|E\\|H\\|F", "G\\|F\\|H\\|E",
                         "H\\|E\\|G\\|F", "H\\|F\\|G\\|E",
                         "EG\\|F\\|H\\|", "EG\\|H\\|F\\|",
                         "EH\\|F\\|G\\|", "EH\\|G\\|F\\|",
                         "FG\\|E\\|H\\|", "FG\\|H\\|E\\|",
                         "FH\\|E\\|G\\|",  "FH\\|G\\|E\\|",
                         "E\\|G\\|FH\\|",  "E\\|H\\|FG\\|",
                         "F\\|G\\|EH\\|",  "F\\|H\\|EG\\|",
                         "G\\|E\\|FH\\|",  "G\\|F\\|EH\\|",
                         "H\\|E\\|FG\\|",  "H\\|F\\|EG\\|")

other_patterns <- c("A\\|BC\\|D\\|",
                    "A\\|BD\\|C\\|",
                    "B\\|AC\\|D\\|",
                    "B\\|AD\\|C\\|",
                    "C\\|BD\\|A\\|",
                    "C\\|AD\\|B\\|",
                    "D\\|BC\\|A\\|",
                    "D\\|AC\\|B\\|",
                    "ACD\\|B\\|\\|",
                    "BCD\\|A\\|\\|",
                    "ABC\\|D\\|\\|",
                    "ABD\\|C\\|\\|",
                    "A\\|BCD\\|\\|",
                    "B\\|ACD\\|\\|",
                    "C\\|ABD\\|\\|",
                    "D\\|ABC\\|\\|",
                    "E\\|FG\\|H\\|",
                    "E\\|FH\\|G\\|",
                    "F\\|EG\\|H\\|",
                    "F\\|EH\\|G\\|",
                    "G\\|FH\\|E\\|",
                    "G\\|EH\\|F\\|",
                    "H\\|FG\\|E\\|",
                    "H\\|EG\\|F\\|",
                    "EGH\\|F\\|\\|",
                    "FGH\\|E\\|\\|",
                    "EFG\\|H\\|\\|",
                    "EFH\\|G\\|\\|",
                    "E\\|FGH\\|\\|",
                    "F\\|EGH\\|\\|",
                    "G\\|EFH\\|\\|",
                    "H\\|EFG\\|\\|")
```

#### directly-provided rankings

```{r}
ranking_direct <- df_judgment  |> 
  filter(sender == "Ranking_Screen") |> 
  select(id, 
         condition, 
         part, 
         order_one_contents_options, 
         order_two_contents_options,
         order_three_contents_options,
         order_four_contents_options)  |>  
  unique() 

sort_letters <- function(strings) {
  sapply(strings, function(x) paste(sort(unlist(strsplit(x, ""))), collapse = ""))
} 
#Alphabetically sort the letters within each cell of the columns "order_one_contents_options", "order_two_contents_options", "order_three_contents_options", "order_four_contents_options" 
#for example convert "DBC" to "BCD"

columns_to_sort <- c("order_one_contents_options", "order_two_contents_options", 
                     "order_three_contents_options", "order_four_contents_options")

ranking_direct[columns_to_sort] <- lapply(ranking_direct[columns_to_sort], sort_letters)

ranking_direct$concatenated_ranking <- apply(ranking_direct[, columns_to_sort], 1, paste, collapse = "|")

ranking_direct <- ranking_direct |>
  mutate(provided_logical = as.integer(sapply(concatenated_ranking, function(x) {
    any(sapply(logical_patterns, function(pattern) grepl(pattern, x)))
  })))  |>
  mutate(provided_stacked = as.integer(sapply(concatenated_ranking, function(x) {
    any(sapply(stacked_patterns, function(pattern) grepl(pattern, x)))
  }))) |>
  mutate(provided_interlaced = as.integer(sapply(concatenated_ranking, function(x) {
    any(sapply(interlaced_patterns, function(pattern) grepl(pattern, x)))
  })))  |>
  mutate(provided_other = as.integer(sapply(concatenated_ranking, function(x) {
    any(sapply(other_patterns, function(pattern) grepl(pattern, x)))
  }))) |>
  select(id, condition, part, provided_ranking = concatenated_ranking, 
         provided_logical, 
         provided_stacked,
         provided_interlaced,
         provided_other) |>
  mutate(provided_ranking_category = case_when(
    provided_logical == 1 ~ "Logical",
    provided_stacked == 1 ~ "Stacked",
    provided_interlaced == 1 ~ "Interlaced",
    provided_other == 1 ~ "Other"
  )) |> 
  mutate(frequency_condition = 
           case_when(condition %in% c("A", "B") ~ "20-learning-trial", 
                     condition %in% c("C", "D") ~ "40-learning-trial",)) |> 
  mutate(check_sum = provided_logical + provided_stacked + provided_interlaced + provided_other)

ranking_direct$provided_ranking_category <- factor(ranking_direct$provided_ranking_category,
                                              levels = c( "Logical","Stacked",
                                                          "Interlaced","Other"),
                                              labels = c("Logical", "Stacked", "Interlaced", "Other"))

```

#### rankings derived from comparisons

```{r}
elicit <- df_judgment |> filter (sender == "show_choice") |> 
  select(id, condition, part, compare1, compare2, response,
         response_action) |> 
  mutate(response = case_when(
    response == "compare1" ~ compare1,
    response == "compare2" ~ compare2,
    TRUE ~ NA
  )) |> 
  group_by(id, condition, part) |> 
  count(response) |> 
  ungroup()

# Define the complete set of response options
response <- c('A', 'B', 'C', 'D',
              'E', 'F', 'G', 'H')

all_options <- expand.grid(id = unique(df_judgment$id), 
                           response= response) 

ranking_comparison <- all_options |> 
  left_join(elicit, by = c("id", "response")) |> #add options that have never been selected in pairwise comparisons to the “response” column.
  mutate(n = ifelse(is.na(n), 0, n)) |>
  group_by(id) |>
  fill(condition, .direction = "downup") |>   
  ungroup() |>                               
  mutate(
    part = case_when(
      response %in% c("A", "B", "C", "D") ~ "mid",
      response %in% c("E", "F", "G", "H") ~ "edge",
      TRUE                                ~ NA_character_
    )) |> 
  group_by(id, part) |> 
  mutate(rank = dense_rank(desc(n))) |>
  select(-n) |>
  pivot_wider(
    names_from  = rank,
    values_from = response
  ) |>
  select(
    id, condition, part,
    order_one   = `1`,
    order_two   = `2`,
    order_three = `3`,
    order_four  = `4`
  ) |> 
  mutate(
    across(
      starts_with("order"),
      ~ map_chr(.x, ~ paste(.x, collapse = ","))  
    )
  ) |> 
  ungroup()

ranking_comparison$concatenated_ranking <- apply(ranking_comparison[, c("order_one", 
                                                "order_two", 
                                                "order_three", 
                                                "order_four")], 1, paste, collapse = "|")

ranking_comparison$concatenated_ranking <- gsub(",| ", "", ranking_comparison$concatenated_ranking) #replace "A, B, C" with "ABC"

ranking_comparison <- ranking_comparison |> 
  mutate(elicited_logical = as.integer(sapply(concatenated_ranking, function(x) {
    any(sapply(logical_patterns, function(pattern) grepl(pattern, x)))
  }))) |> 
  mutate(elicited_stacked = as.integer(sapply(concatenated_ranking, function(x) {
    any(sapply(stacked_patterns, function(pattern) grepl(pattern, x)))
  })))|> 
  mutate(elicited_interlaced = as.integer(sapply(concatenated_ranking, function(x) {
    any(sapply(interlaced_patterns, function(pattern) grepl(pattern, x)))
  }))) |> 
  mutate(elicited_other = as.integer(sapply(concatenated_ranking, function(x) {
    any(sapply(other_patterns, function(pattern) grepl(pattern, x)))
  }))) |> 
  mutate(check_sum = elicited_logical + elicited_stacked + elicited_interlaced + elicited_other) |> 
  mutate(frequency_condition = case_when(
    condition %in% c("A", "B") ~ "20-learning-trial",
    condition %in% c("C", "D") ~ "40-learning-trial",
    TRUE ~ NA_character_  
  )) |>
  select(id, condition, frequency_condition, part, 
         elicited_ranking = concatenated_ranking,
         elicited_logical, 
         elicited_stacked,
         elicited_interlaced,
         elicited_other,
         check_sum) |>
  mutate(elicited_ranking_category = case_when(
    elicited_logical == 1 ~ "Logical",
    elicited_stacked == 1 ~ "Stacked",
    elicited_interlaced == 1 ~ "Interlaced",
    elicited_other == 1 ~ "Other"
  )) 

ranking_comparison$elicited_ranking_category <- factor(ranking_comparison$elicited_ranking_category,
                                                   levels = c( "Logical","Stacked",
                                                               "Interlaced","Other"),
                                                   labels = c("Logical", "Stacked", 
                                                              "Interlaced", "Other"))

```

### Plot

```{r}
ranking_direct_subsample <- ranking_direct |> 
  filter(id %in% subsample_ids)

ranking_comparison_subsample <- ranking_comparison |> 
  filter(id %in% subsample_ids)

prep <- function(df, catcol, source, sample) {
  df |> 
    transmute(id, part, frequency_condition,
              ranking_category = {{ catcol }},
              ranking_source = source,
              sample = sample)
}

ranking_all <- bind_rows(
  prep(ranking_direct,               provided_ranking_category, "Directly provided", "Whole sample"),
  prep(ranking_comparison,           elicited_ranking_category, "Pairwise-derived",  "Whole sample"),
  prep(ranking_direct_subsample,     provided_ranking_category, "Directly provided", "Subsample"),
  prep(ranking_comparison_subsample, elicited_ranking_category, "Pairwise-derived",  "Subsample")
) |> 
  mutate(
    sample              = factor(sample, levels = c("Whole sample", "Subsample")),
    ranking_source      = factor(ranking_source,
                                 levels = c("Directly provided", "Pairwise-derived")),
    frequency_condition = factor(frequency_condition,
                                 levels = c("40-learning-trial", "20-learning-trial"),
                                 labels = c("40 trials", "20 trials")),
    part                = factor(part, levels = c("edge", "mid"),
                                 labels = c("Edge", "Mid")),
    # reversed so Logical anchors the bottom of each stack
    ranking_category    = factor(ranking_category,
                                 levels = c("Other", "Interlaced", "Stacked", "Logical"))
  ) |> 
  # row label carrying N, ordered to follow `sample`
  group_by(sample) |> 
  mutate(sample_lab = paste0(sample, "\nN = ", n_distinct(id))) |> 
  ungroup() |> 
  mutate(sample_lab = fct_reorder(sample_lab, as.integer(sample)))

theme_apa <- theme_classic(base_size = 11) +
  theme(
    axis.text        = element_text(size = 10, colour = "black"),
    axis.line        = element_line(colour = "black", linewidth = 0.4),
    axis.ticks       = element_line(colour = "black", linewidth = 0.4),
    strip.background = element_rect(fill = NA, colour = "black", linewidth = 0.4),
    strip.text.x     = element_text(size = 10, margin = margin(3, 3, 3, 3)),
    strip.text.y     = element_text(size = 10, angle = -90,
                                    margin = margin(3, 3, 3, 3)),
    legend.position  = "bottom",
    legend.key.size  = unit(0.8, "lines"),
    panel.spacing.x  = unit(0, "lines"),
    panel.spacing.y  = unit(0.7, "lines")
  )

p_apa <- ggplot(ranking_all, aes(x = part, fill = ranking_category)) +
  geom_bar(position = "fill", width = 0.55, colour = "white", linewidth = 0.2) +
  facet_nested(sample_lab ~ ranking_source + frequency_condition) +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.02))) +
  scale_fill_manual(
    name   = "Ranking type",
    values = c(Logical = "#009E73", Stacked = "#F0E442",
               Interlaced = "#CC79A7", Other = "#0072B2"),
    breaks = c("Logical", "Stacked", "Interlaced", "Other")
  ) +
  guides(fill = guide_legend(reverse = TRUE, nrow = 1)) +
  labs(x = "Event set type", y = "Percentage of participants") +
  theme_apa

p_apa

ggsave("judgment.pdf", p_apa, width = 7, height = 5.5, units = "in")
```

### MPT analysis

```{r}
# Quick checks: every ranking falls in exactly one category,
# and there is exactly one ranking per participant and event set
stopifnot(all(ranking_direct$check_sum == 1))
stopifnot(all(ranking_comparison$check_sum == 1))
stopifnot(!any(duplicated(ranking_direct[, c("id", "part")])))
stopifnot(!any(duplicated(ranking_comparison[, c("id", "part")])))

count_cells <- function(df, logical, other, stacked, interlaced) {
  df |>
    group_by(frequency_condition, part) |>
    summarise(logical    = sum({{ logical }}),
              other      = sum({{ other }}),
              stacked    = sum({{ stacked }}),
              interlaced = sum({{ interlaced }}),
              n          = n(),
              .groups    = "drop")
}

counts_direct <- count_cells(ranking_direct,
                             provided_logical, provided_other,
                             provided_stacked, provided_interlaced)

counts_comparison <- count_cells(ranking_comparison,
                                 elicited_logical, elicited_other,
                                 elicited_stacked, elicited_interlaced)

counts_direct
counts_comparison

# prepare data in the format that MPTinR wants 
one_tree <- function(counts, cond, set) {
  row <- counts |> filter(frequency_condition == cond, part == set)
  c(row$logical, row$other, row$stacked, row$interlaced)
}

data_vector <- function(counts) {
  c(one_tree(counts, "20-learning-trial", "edge"),   # tree 1: edge20
    one_tree(counts, "20-learning-trial", "mid"),    # tree 2: mid20
    one_tree(counts, "40-learning-trial", "edge"),   # tree 3: edge40
    one_tree(counts, "40-learning-trial", "mid"))    # tree 4: mid40
}

data_direct     <- data_vector(counts_direct)
data_comparison <- data_vector(counts_comparison)
data_direct
data_comparison
```

```{r}
# ---------------------- The preregistered four-tree model
model_unrestricted <- c(
  # tree 1: edge event set, 20 learning trials
  "le20",
  "(1-le20)*oe20",
  "(1-le20)*(1-oe20)*se20",
  "(1-le20)*(1-oe20)*(1-se20)",
  "",
  # tree 2: mid event set, 20 learning trials
  "lm20",
  "(1-lm20)*om20",
  "(1-lm20)*(1-om20)*sm20",
  "(1-lm20)*(1-om20)*(1-sm20)",
  "",
  # tree 3: edge event set, 40 learning trials
  "le40",
  "(1-le40)*oe40",
  "(1-le40)*(1-oe40)*se40",
  "(1-le40)*(1-oe40)*(1-se40)",
  "",
  # tree 4: mid event set, 40 learning trials
  "lm40",
  "(1-lm40)*om40",
  "(1-lm40)*(1-om40)*sm40",
  "(1-lm40)*(1-om40)*(1-sm40)"
)

model_file <- tempfile(fileext = ".model")
writeLines(model_unrestricted, model_file)

check.mpt(model_file)   # expect: 4 trees, 16 categories, 12 parameters

fit_direct     <- fit.mpt(data_direct, model_file)
fit_direct

fit_comparison <- fit.mpt(data_comparison, model_file)
fit_comparison
```

```{r}
# Because the model is saturated, every parameter is simply an observed
# proportion:
#   l = logical / n
#   O = other   / (n - logical)
#   S = stacked / (n - logical - other)
# so exact (Clopper-Pearson) 95% CIs come from binom.test().
# MPTinR's own CIs are Wald intervals and go below 0 for small parameters
# (e.g. oe20, oe40)

exact_ci_table <- function(counts) {
  out <- data.frame()
  for (i in 1:nrow(counts)) {
    r      <- counts[i, ]
    suffix <- paste0(ifelse(r$part == "edge", "e", "m"),
                     ifelse(r$frequency_condition == "20-learning-trial", "20", "40"))
    x <- c(r$logical, r$other,          r$stacked)
    n <- c(r$n,       r$n - r$logical,  r$n - r$logical - r$other)
    for (j in 1:3) {
      ci  <- binom.test(x[j], n[j])$conf.int
      out <- rbind(out, data.frame(
        parameter   = paste0(c("l", "o", "s")[j], suffix),
        numerator   = x[j],
        denominator = n[j],
        estimate    = x[j] / n[j],
        ci_lower    = ci[1],
        ci_upper    = ci[2]))
    }
  }
  out |> mutate(across(c(estimate, ci_lower, ci_upper), ~ round(.x, 3)))
}

ci_direct     <- exact_ci_table(counts_direct)
ci_comparison <- exact_ci_table(counts_comparison)

ci_direct
ci_comparison
```

```{r}
# Sanity check: the proportions must equal MPTinR's estimates. If this fails,
stopifnot(max(abs(ci_direct$estimate -
                    fit_direct$parameters[ci_direct$parameter, "estimates"])) < 0.001)
stopifnot(max(abs(ci_comparison$estimate -
                    fit_comparison$parameters[ci_comparison$parameter, "estimates"])) < 0.001)
```

```{r}
# ----------------------- Preregistered hypotheses: likelihood-ratio tests
# ---------- main hypotheses:c("le20", "om20", "sm20", "le40", "om40", "sm40") > c("lm20", "oe20", "se20", "lm40", "oe40", "se40")
# ---------- secondary research question: c("le20", "lm20", "om40", "sm40") > c("le40", "lm40", "om20", "sm20")
# Each hypothesis says "larger > smaller". We fit a model in which the two
# parameters are forced to be EQUAL and compare it with the unrestricted model:
#   G2 difference (df = 1) large  ->  the two parameters differ.
# Then we check whether the difference goes in the predicted direction.

hypotheses <- data.frame(
  rq      = c(rep("Main", 6), rep("RQ b", 4)),
  larger  = c("le20", "om20", "sm20", "le40", "om40", "sm40",
              "le20", "lm20", "om40", "sm40"),
  smaller = c("lm20", "oe20", "se20", "lm40", "oe40", "se40",
              "le40", "lm40", "om20", "sm20")
)

lrt_table <- function(data, fit_unrestricted) {
  out <- data.frame()
  for (i in 1:nrow(hypotheses)) {
    larger  <- hypotheses$larger[i]
    smaller <- hypotheses$smaller[i]
    
    fit_restricted <- fit.mpt(data, model_file,
                              restrictions.filename = list(paste(smaller, "=", larger)),
                              show.messages = FALSE)
    
    G2 <- fit_restricted$goodness.of.fit$G.Squared - fit_unrestricted$goodness.of.fit$G.Squared
    df <- fit_restricted$goodness.of.fit$df        - fit_unrestricted$goodness.of.fit$df
    
    est_larger  <- fit_unrestricted$parameters[larger,  "estimates"]
    est_smaller <- fit_unrestricted$parameters[smaller, "estimates"]
    
    out <- rbind(out, data.frame(
      rq           = hypotheses$rq[i],
      hypothesis   = paste(larger, ">", smaller),
      est_larger   = est_larger,
      est_smaller  = est_smaller,
      direction_ok = est_larger > est_smaller,
      G2           = G2,
      df           = df,
      p            = pchisq(G2, df, lower.tail = FALSE),
      # positive dAIC / dBIC = the unrestricted model (parameters differ) is preferred
      dAIC         = fit_restricted$information.criteria$AIC - fit_unrestricted$information.criteria$AIC,
      dBIC         = fit_restricted$information.criteria$BIC - fit_unrestricted$information.criteria$BIC))
  }
  rownames(out) <- NULL
  out |> mutate(across(c(est_larger, est_smaller, G2, dAIC, dBIC), ~ round(.x, 3)),
                p = round(p, 4))
}

lrt_direct     <- lrt_table(data_direct,     fit_direct)
lrt_comparison <- lrt_table(data_comparison, fit_comparison)

lrt_direct
lrt_comparison
```

```{r}
# 5. RQ: agreement between the two elicitation methods (Cohen's kappa)

# Sort the letters inside each tie slot ("A|DC|B|" -> "A|CD|B|") so that the
# two methods write the same ranking the same way before we compare them.
sort_slots <- function(ranking) {
  sapply(strsplit(ranking, "\\|"), function(slots) {
    paste(sapply(slots, function(s) paste(sort(strsplit(s, "")[[1]]), collapse = "")),
          collapse = "|")
  })
}

agreement <- ranking_direct |>
  select(id, part, frequency_condition, provided_ranking, provided_ranking_category) |>
  inner_join(ranking_comparison |>
               select(id, part, elicited_ranking, elicited_ranking_category),
             by = c("id", "part")) |>
  mutate(provided_ranking = sort_slots(provided_ranking),
         elicited_ranking = sort_slots(elicited_ranking))

stopifnot(nrow(agreement) == nrow(ranking_direct))

# preregistered: agreement on the exact ranking
kappa_ranking <- kappa2(agreement[, c("provided_ranking", "elicited_ranking")])
kappa_ranking
mean(agreement$provided_ranking == agreement$elicited_ranking)   # proportion identical

# supplementary: agreement on the ranking category
kappa_category <- kappa2(agreement[, c("provided_ranking_category", "elicited_ranking_category")])
kappa_category
mean(agreement$provided_ranking_category == agreement$elicited_ranking_category)
```

##### Subsample for robustness analyses 

```{r}
length(subsample_ids)     # expect 310

ranking_direct_sub     <- ranking_direct     |> filter(id %in% subsample_ids)
ranking_comparison_sub <- ranking_comparison |> filter(id %in% subsample_ids)

# every subsample id should have one ranking per event set in both files
stopifnot(nrow(ranking_direct_sub)     == 2 * length(subsample_ids))
stopifnot(nrow(ranking_comparison_sub) == 2 * length(subsample_ids))

counts_direct_sub <- count_cells(ranking_direct_sub,
                                 provided_logical, provided_other,
                                 provided_stacked, provided_interlaced)

counts_comparison_sub <- count_cells(ranking_comparison_sub,
                                     elicited_logical, elicited_other,
                                     elicited_stacked, elicited_interlaced)

counts_direct_sub       # expect n = 147 (20 trials) and 163 (40 trials)
counts_comparison_sub

data_direct_sub     <- data_vector(counts_direct_sub)
data_comparison_sub <- data_vector(counts_comparison_sub)

data_direct_sub
data_comparison_sub

# some cells are empty 
#"other" in edge20 for the direct rankings
# "stacked" in edge20 for the pairwise-derived rankings

fit_direct_sub     <- fit.mpt(data_direct_sub,     model_file, show.messages = FALSE)
fit_comparison_sub <- fit.mpt(data_comparison_sub, model_file, show.messages = FALSE)

fit_direct_sub$goodness.of.fit          # G2 = 0, df = 0 (saturated)
fit_comparison_sub$goodness.of.fit

ci_direct_sub     <- exact_ci_table(counts_direct_sub)
ci_comparison_sub <- exact_ci_table(counts_comparison_sub)

# An empty cell gives an estimate of exactly 0. The exact CI handles this
# (e.g. [.000, .08]); MPTinR's Wald CI would be [0, 0], which is meaningless.
ci_direct_sub
ci_comparison_sub

stopifnot(max(abs(ci_direct_sub$estimate -
                    fit_direct_sub$parameters[ci_direct_sub$parameter, "estimates"])) < 0.001)
stopifnot(max(abs(ci_comparison_sub$estimate -
                    fit_comparison_sub$parameters[ci_comparison_sub$parameter, "estimates"])) < 0.001)

```

```{r}
# -------------------------------- Preregistered hypotheses: likelihood-ratio tests

lrt_direct_sub     <- lrt_table(data_direct_sub,     fit_direct_sub)
lrt_comparison_sub <- lrt_table(data_comparison_sub, fit_comparison_sub)

lrt_direct_sub
lrt_comparison_sub

# CAUTION: tests involving a parameter estimated from a very small or empty
# cell (om20 > oe20 in the direct rankings, sm20 > se20 in the pairwise
# rankings) rely on the chi-square approximation with sparse data. If one of
# these is close to p = .05, check it with the bootstrap in block F of
# 02_judgment_additional.R, e.g.:
#   boot_lrt_p(data_direct_sub,     "om20", "oe20")
#   boot_lrt_p(data_comparison_sub, "sm20", "se20")
```

```{r}
# ------------------------- Cohen's kappa

agreement_sub <- ranking_direct_sub |>
  select(id, part, frequency_condition, provided_ranking, provided_ranking_category) |>
  inner_join(ranking_comparison_sub |>
               select(id, part, elicited_ranking, elicited_ranking_category),
             by = c("id", "part")) |>
  mutate(provided_ranking = sort_slots(provided_ranking),
         elicited_ranking = sort_slots(elicited_ranking))

stopifnot(nrow(agreement_sub) == nrow(ranking_direct_sub))

kappa_ranking_sub <- kappa2(agreement_sub[, c("provided_ranking", "elicited_ranking")])
kappa_ranking_sub
mean(agreement_sub$provided_ranking == agreement_sub$elicited_ranking)

kappa_category_sub <- kappa2(agreement_sub[, c("provided_ranking_category", "elicited_ranking_category")])
kappa_category_sub
mean(agreement_sub$provided_ranking_category == agreement_sub$elicited_ranking_category)
```

```{r}
# ------------ Side-by-side comparison with the whole sample
# Does any conclusion change when only satisfactory learners are analysed?

# ---- Parameter estimates ------------------------------------------------------
compare_estimates <- function(ci_whole, ci_sub) {
  ci_whole |>
    select(parameter, whole = estimate) |>
    left_join(ci_sub |> select(parameter, sub = estimate,
                               sub_lower = ci_lower, sub_upper = ci_upper),
              by = "parameter") |>
    mutate(change = round(sub - whole, 3))
}

estimates_compared_direct     <- compare_estimates(ci_direct,     ci_direct_sub)
estimates_compared_comparison <- compare_estimates(ci_comparison, ci_comparison_sub)

estimates_compared_direct
estimates_compared_comparison

# Summary of how much the estimates moved (for the robustness paragraph)
all_changes <- abs(c(estimates_compared_direct$change,
                     estimates_compared_comparison$change))
data.frame(median_abs_change = median(all_changes),
           max_abs_change    = max(all_changes))


# ---- Hypothesis tests ---------------------------------------------------------
# A conclusion counts as "changed" only if
#   (a) the test is significant in one sample but not in the other, or
#   (b) the test is significant in both samples but in opposite directions.
# The direction of a NON-significant difference is not interpreted, so a sign
# flip between two non-significant tests (e.g. .759 vs .778 -> .776 vs .767)
# does not count as a change.

compare_tests <- function(lrt_whole, lrt_sub) {
  lrt_whole |>
    select(rq, hypothesis,
           direction_whole = direction_ok, G2_whole = G2, p_whole = p) |>
    left_join(lrt_sub |>
                select(hypothesis,
                       direction_sub = direction_ok, G2_sub = G2, p_sub = p),
              by = "hypothesis") |>
    mutate(sig_whole       = p_whole < .05,
           sig_sub         = p_sub   < .05,
           same_conclusion = (sig_whole == sig_sub) &
             (!sig_whole | direction_whole == direction_sub))
}

tests_compared_direct     <- compare_tests(lrt_direct,     lrt_direct_sub)
tests_compared_comparison <- compare_tests(lrt_comparison, lrt_comparison_sub)

tests_compared_direct
tests_compared_comparison

# Which conclusions change? 
# -> only om40 > oe40, pairwise-derived)
tests_compared_direct     |> filter(!same_conclusion)
tests_compared_comparison |> filter(!same_conclusion)

# ---- Cohen's kappa ------------------------------------------------------------
data.frame(sample         = c("Whole", "Subsample"),
           n_rankings     = c(nrow(agreement), nrow(agreement_sub)),
           kappa_ranking  = round(c(kappa_ranking$value,  kappa_ranking_sub$value),  3),
           kappa_category = round(c(kappa_category$value, kappa_category_sub$value), 3))

```
